---
title: "Final Project: Occcupations"
names: "Lily Jakielaszek, Allison Denenberg, Ayoolwatomiwa Ogunrombi"
output: html_notebook
---

### Load the Libraries 
```{r}
library(DataComputing)
library(lubridate)
library(tidyverse)
```

### Load the Main Dataset from Outside Source
```{r}
OccupationProjection <- read.csv("https://query.data.world/s/hfgf3loqzzbk34aytghvne5sxg3ljr", header=TRUE, stringsAsFactors=FALSE);
OccupationProjection
```
### Load Dataset from R package 
```{r}
salaries_by_region = readr::read_csv("salaries-by-region.csv")
salaries_by_college_type = readr::read_csv("salaries-by-college-type.csv")
degrees_that_pay_back = readr::read_csv("degrees-that-pay-back.csv")
```
### Research Question
What occupations and majors by state will have the highest salary and have the greatest future outlook?

### Becoming Acquainted with the Dataset 
1. Where did you find them?
2. Who collected/maintains them?
3. When & Why were they originally collected?
4. What does a case represent in each data source, and how many total cases are available?
5. What are some of the variables that you plan to use
(The R Notebook should include R code chunks & output as well as narrative (bullet list is fine at this stage) as you answer these questions)