---
title: "Final Project: Occcupations"
names: "Lily Jakielaszek, Allison Denenberg, Ayoolwatomiwa Ogunrombi"
output: html_notebook
---

### Load the Libraries 
```{r}
library(DataComputing)
library(lubridate)
library(tidyverse)
library(rvest)
library(mosaic)
```

### Load the Main Dataset from Outside Source
```{r}
OccupationProjection <- read.csv("https://query.data.world/s/hfgf3loqzzbk34aytghvne5sxg3ljr", header=TRUE, stringsAsFactors=FALSE);
OccupationProjection<-OccupationProjection%>%
  filter(Occupation.Name!="Total, All Occupations")
```
### Load Dataset from R package 
```{r}
salaries_by_region = readr::read_csv("salaries-by-region.csv")
salaries_by_college_type = readr::read_csv("salaries-by-college-type.csv")
degrees_that_pay_back = readr::read_csv("degrees-that-pay-back.csv")
```
### Research Question
What occupations and majors by state will have the highest salary and have the greatest future outlook?

### Quick Look at the Datasets
```{r}
nrow(OccupationProjection)
ncol(OccupationProjection)
```

```{r}
nrow(salaries_by_region)
ncol(salaries_by_region)
```

```{r}
nrow(salaries_by_college_type)
ncol(salaries_by_college_type)
```

```{r}
nrow(degrees_that_pay_back)
ncol(degrees_that_pay_back)
```


### Becoming Acquainted with the Dataset 
1. Where did you find them?

The first dataset was found on Data.World, the other three datasets were found on Kaggle. 

2. Who collected/maintains them?

The data from Data.World swas collected and updated by J. Albert Bowden II 
The data from Kaggle says that all data was obtained from the Wall Street Journal based on data from Payscale, Inc 3 years ago

3. When & Why were they originally collected?

Both datasets were created 3 years ago and both had updates only 3 years ago as well.
For the Data.World dataset, it was made to informed career decisions 
For the Kaggle salaries by region dataset, it was made to look at schools and regions salaries by mid-career.
For the Kaggle salaries by college_type dataset, it was made to compare rankings of type of colleges you attended (ivy league, engeenering school)
For the Kaggle degrees that pay back dataset, it was made to define which majors make the most money 

4. What does a case represent in each data source, and how many total cases are available?

1. 
```{r}
glimpse(OccupationProjection)
```
A case represents an area such as nation (United States) or state and an occupation name. There are 34,982 cases.
2.
```{r}
glimpse(salaries_by_region)
```
A case represents a school name and region. There are 320 cases.
3.
```{r}
glimpse(salaries_by_college_type)
```
A case represents a school name and school type. There are 269 cases.
4. 
```{r}
glimpse(degrees_that_pay_back)
```
A case represents an undergraduate major. There are 50 cases. 

5. What are some of the variables that you plan to use

We plan to use School Name, School Type, Mid-Career Median Salary, Starting Median Salary, Percent change from Starting to Mid-Career Salary, Region, Area.name, Occupation.name, Base.Year, Projected.Year, Average.Annual.Openings, Percent.Change 

We plan to connect some of these datasets by the State name, as well as use some datasets alone. 

### Explore intuition related to the research question
1. Create some informative plots and summary statistics 

a. 
```{r}
degrees_that_pay_back%>%
  ggplot(aes(x=`Undergraduate Major`,y=`Percent change from Starting to Mid-Career Salary`))+geom_point()+theme_bw()
```
```{r}
OccupationProjection<-OccupationProjection%>%
  filter(Average.Annual.Openings>=20000)
OccupationProjection
```

```{r}
ggplot(data=OccupationProjection,aes(x=Occupation.Name,y=Average.Annual.Openings ))+geom_bar(stat='identity',position='stack', width=.9)+theme_bw()
```





