---
title: "FinalProjectData"
authors: "Lily Jakielaszek, Allison Denenberg, Ayoolwatomiwa Ogunrombi"
output: html_notebook
---

### Load the Libraries and Datasets
```{r}
# front matter
rm(list = ls())
library(DataComputing)
library(lubridate)
library(tidyverse)
library(rvest)
library(mosaic)
library(party)
OccupationProjection <- read.csv("https://query.data.world/s/hfgf3loqzzbk34aytghvne5sxg3ljr", header=TRUE, stringsAsFactors=FALSE);

OccupationProjection<-OccupationProjection%>%
  filter(Occupation.Name!="Total, All Occupations")

salaries_by_region = readr::read_csv("salaries-by-region.csv")

salaries_by_college_type = readr::read_csv("salaries-by-college-type.csv")

degrees_that_pay_back = readr::read_csv("degrees-that-pay-back.csv")
```
# Research Question

What school, major, and occupation should a prospective college student choose (what is the game plan)?

## Make New Datasets

These datasets will be the ones that will be used to perform statistics on and joins on. They are renamed from the original datasets and include all the same information except they have R acceptable variable names and the correct type for the their output. 

```{r}
SalariesByRegion <-
  salaries_by_region %>%
  mutate(School_Name = `School Name`,
         Mid_Career_Median_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career Median Salary`)),
         Starting_Median_Salary = as.numeric(gsub('\\$|,', '', `Starting Median Salary`)),
         Mid_Career_10th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 10th Percentile Salary`)),
         Mid_Career_25th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 25th Percentile Salary`)),
         Mid_Career_75th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 75th Percentile Salary`)),
         Mid_Career_90th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 90th Percentile Salary`)),
         Difference = Mid_Career_Median_Salary - Starting_Median_Salary) 
SalariesByRegion <- 
  SalariesByRegion %>%
  select(School_Name, Region, Mid_Career_Median_Salary, Starting_Median_Salary, Mid_Career_10th_Percentile_Salary, Mid_Career_25th_Percentile_Salary, Mid_Career_75th_Percentile_Salary, Mid_Career_90th_Percentile_Salary, Difference)
SalariesByRegion
```
```{r}
SalariesByCollegeType <-
  salaries_by_college_type %>%
  mutate(School_Name = `School Name`,
         School_Type = `School Type`,
         Mid_Career_Median_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career Median Salary`)),
         Starting_Median_Salary = as.numeric(gsub('\\$|,', '', `Starting Median Salary`)),
         Mid_Career_10th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 10th Percentile Salary`)),
         Mid_Career_25th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 25th Percentile Salary`)),
         Mid_Career_75th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 75th Percentile Salary`)),
         Mid_Career_90th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 90th Percentile Salary`)),
         Difference = Mid_Career_Median_Salary - Starting_Median_Salary)
SalariesByCollegeType <-
  SalariesByCollegeType %>%
  select(School_Name, School_Type, Starting_Median_Salary, Mid_Career_Median_Salary, Mid_Career_10th_Percentile_Salary, Mid_Career_25th_Percentile_Salary, Mid_Career_75th_Percentile_Salary, Mid_Career_90th_Percentile_Salary, Difference)
SalariesByCollegeType
```

```{r}
DegreesThatPayBack <-
  degrees_that_pay_back %>%
  mutate(Undergraduate_Major = `Undergraduate Major`,
         Mid_Career_Median_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career Median Salary`)),
         Starting_Median_Salary = as.numeric(gsub('\\$|,', '', `Starting Median Salary`)),
         Mid_Career_10th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 10th Percentile Salary`)),
         Mid_Career_25th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 25th Percentile Salary`)),
         Mid_Career_75th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 75th Percentile Salary`)),
         Mid_Career_90th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 90th Percentile Salary`)),
         Percent_Change_From_Starting_To_MidCareer_Salary = `Percent change from Starting to Mid-Career Salary`,
         Difference = Mid_Career_Median_Salary - Starting_Median_Salary)
DegreesThatPayBack <-
  DegreesThatPayBack %>%
  select(Undergraduate_Major, Starting_Median_Salary, Mid_Career_Median_Salary, Mid_Career_10th_Percentile_Salary, Mid_Career_25th_Percentile_Salary, Mid_Career_75th_Percentile_Salary, Mid_Career_90th_Percentile_Salary, Percent_Change_From_Starting_To_MidCareer_Salary, Difference)
DegreesThatPayBack
```
```{r}
OccupationsProjection <-
  OccupationProjection %>%
  mutate(State_FIPS = State.FIPS,
         Region = Area.Name,
         Occupation_Code = Occupation.Code,
         Occupation_Name = Occupation.Name,
         Base_Year = Base.Year,
         Base = Base,
         Projected_Year = Projected.Year,
         Projection= Projection,
         Percent_Change = Percent.Change,
         Average_Annual_Openings = Average.Annual.Openings) 
OccupationsProjection <-
  OccupationsProjection %>%
  select(State_FIPS, Region, Occupation_Code, Occupation_Name, Base_Year, Base, Projected_Year, Projection, Percent_Change, Average_Annual_Openings) 
OccupationsProjection
```
### Inspecting the Revised Datasets 
## OccupationProjection Inspection
```{r}
str(OccupationProjection)
head(OccupationProjection)
tail(OccupationProjection)
nrow(OccupationProjection)
```

## DegreesThatPayBack Inspection
```{r}
str(DegreesThatPayBack)
head(DegreesThatPayBack)
tail(DegreesThatPayBack)
nrow(DegreesThatPayBack)
```

## SalariesByCollegeType Inspection
```{r}
str(SalariesByCollegeType)
head(SalariesByCollegeType)
tail(SalariesByCollegeType)
nrow(SalariesByCollegeType)
```

## SalariesByRegion Inspection
```{r}
str(SalariesByRegion)
head(SalariesByRegion)
tail(SalariesByRegion)
nrow(SalariesByRegion)
```

### First Step in Game Plan: Find the Schools with the Highest Salaries

Join the datasets together to match up the school names. 
```{r}
SchoolSalaries <-
  SalariesByCollegeType %>%
  full_join(SalariesByRegion)
head(SchoolSalaries)
```

We are looking at side by side box plots of each school type and comparing their starting median salaries.
```{r}
ggplot(SchoolSalaries, aes(x = School_Type, y = Starting_Median_Salary)) +
  geom_boxplot() +
  xlab("School Type") +
  ylab("Starting Median Salary") +
  ggtitle("Starting Median Salary Based on School Type")
```

Now we will observe the top ten schools with the highest starting median salaries.
```{r}
SchoolSalaries %>%
  select(School_Name, School_Type, Region, Starting_Median_Salary) %>%
  arrange(desc(Starting_Median_Salary)) %>% 
  head(10)
```

Now we will find the top ten schools with the highest mid-career median salaries.
```{r}
MidCareerMedianSalaries <-
  SchoolSalaries %>%
  select(School_Name, School_Type, Region, Mid_Career_Median_Salary) %>%
  arrange(desc(Mid_Career_Median_Salary)) %>% 
  head(10)
MidCareerMedianSalaries
```
At this point we can see that California and NorthEastern schools are in the top ten consistenly. We can also see that the top school types are Ivy League and Engineering schools. 

Now we will combine these results to find the best school for percent change from starting to mid-career median salary which will show the best school for salary growth over time.
```{r}
percentChange <-
  SchoolSalaries %>%
  mutate(Percent_Change = ((Mid_Career_Median_Salary - Starting_Median_Salary)/Starting_Median_Salary) * 100) %>%
  select(School_Name, School_Type, Percent_Change) %>%
  arrange(desc(Percent_Change)) %>%
  head(50)
percentChange
```

## Visualizations on Schools and Salaries 

```{r}
SchoolSalariesGraphic <-
  percentChange%>%
  ggplot(aes(x = reorder(School_Name, -Percent_Change), y = Percent_Change, color = School_Type)) + 
  geom_point() +
  xlab("School Name") +
  ylab("Percent Change") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

SchoolSalariesGraphic

```

```{r}
SchoolSalaries %>% # This code will find the mean for mid career salary
  summarise(mean = sum(Mid_Career_Median_Salary)/339)
MidCareerAvgMedianSalary <-
  SchoolSalaries %>%
  na.omit() %>%
  mutate(Above_Avg_MidCareer_Median_Salary = ifelse(Mid_Career_Median_Salary <= 86076.11, "yes", "no")) # This code uses the mean to indicate an above or below for each school
MidCareerAvgMedianSalary
```

Now we will make a tree to demostrate the importance of school type and region in determining whether a school will have an above average mid career median salary.

```{r}
TreeForMidCareer <- 
  MidCareerAvgMedianSalary %>%
  select(Above_Avg_MidCareer_Median_Salary, Starting_Median_Salary,Mid_Career_Median_Salary,School_Type, Region) 
TreeForMidCareer<-TreeForMidCareer%>%mutate(School_Type=as.factor(School_Type), Region=as.factor(Region))
mod1 <- party::ctree(Above_Avg_MidCareer_Median_Salary=="yes"~ School_Type+Region, data = TreeForMidCareer)
plot(mod1,type="simple")
```

According to this decision tree, the probability of having an above median mid career salary if you go to a state school leads to high chance of a high salary. This is a little shocking since Engineering and Ivy League Schools are grouped with Liberal Arts and Party Schools. We therefore believe that the size of our data-set is affecting our results and have chosen to ignore this conclusion. However, according to the desicion tree, If you go to a State School in the Midwestern, Southern or Western regions, you have a high probablity of 0.745 of getting an above median salary mid career. If you go to a Engineering or Ivy League Schools or Liberal Arts or Party Schools in the Southern or Western regions, you have a high probability of 0.474 of getting an above median salary mid career. This shows that the Southern and Western regions are the best indicators in the tree as leading to a higher salary, while the California Region leads to a low chance of a high salary no matter what school type you go to.


We will do the same thing as up above but now we will look at starting median salary.

```{r}
SchoolSalaries %>% # This code will find the mean for starting salary
  summarise(mean = sum(Starting_Median_Salary)/349)
StartingAvgMedianSalary <-
  SchoolSalaries %>%
  mutate(Above_Avg_Starting_Median_Salary = ifelse(Starting_Median_Salary <= 46093.41, "yes", "no"))
StartingAvgMedianSalary
```

Now we will make a tree to demonstrate the importance of school type and region in determining whether a school will have an above average starting median salary.

```{r}

CompleteCases <- 
  StartingAvgMedianSalary %>%
  select(Above_Avg_Starting_Median_Salary, Starting_Median_Salary,Mid_Career_Median_Salary,School_Type, Region,School_Name) %>%
  na.omit() 
CompleteCases<-CompleteCases%>%mutate(School_Type=as.factor(School_Type), Region=as.factor(Region),School_Name=as.factor(School_Name))
mod1 <- party::ctree(Above_Avg_Starting_Median_Salary=="yes"~ School_Type+Region+School_Name, data = CompleteCases)
plot(mod1,type="simple")

```

According to this decision tree, the probability of having an above median salary if you go to an Ivy-League or Engineering School is zero, which we know cannot be true. We therefore believe that the size of our data-set is affecting our results and have chosen to ignore this conclusion. However, according to the desicion tree, If you go to a School in the Midwestern, Southern or Western regions, you have a high probablity of 0.753 of getting an above median salary.

### Conclusions on Schools and Salaries 

We can conclude that the school to pick if you want the highest starting median salary is California Institute of Technology (CIT) with a salary amount of $75,500, the school to pick if you want the highest mid-career median salary is Dartmouth College with a salary amount of $134,000, and the school to pick if you want the best salary growth over time is DartMouth College with a salary percent growth from starting to mid-career median salary of 131.0345%. 

# Second Step in Game Plan: Finding the Majors with the Highest Salaries 

We will be using the DegreesThatPayBack dataset to answer this part of the game plan.

First we will look at the top ten majors with the best starting median salary.

```{r}
head(DegreesThatPayBack)
DegreesThatPayBack %>%
  select(Undergraduate_Major, Starting_Median_Salary) %>%
  arrange(desc(Starting_Median_Salary)) %>%
  head(10)
```
From this we can see that the best major for starting median salary is a Physician Assistant.

Now we will look at the top ten majors with the best mid-career median salary.
```{r}
DegreesThatPayBack %>%
  select(Undergraduate_Major, Mid_Career_Median_Salary) %>%
  arrange(desc(Mid_Career_Median_Salary)) %>%
  head(10)
```
From this we can see that the best major for mid-career median salary is Chemical Engineering.

Now we will look at the top ten majors with the best percentage change from starting to mid-career median salary to see which major offers the best growth in salary.
```{r}
DegreesThatPayBack <-
  DegreesThatPayBack%>%
  select(Undergraduate_Major, Percent_Change_From_Starting_To_MidCareer_Salary) %>%
  arrange(desc(Percent_Change_From_Starting_To_MidCareer_Salary)) %>%
  head(50)
DegreesThatPayBack
```
From this we can see that Math is best major for average salary growth.

## Visualization of Majors and Salaries

```{r}
DegreesThatPayBack %>%
  ggplot(aes(x = reorder(Undergraduate_Major, -Percent_Change_From_Starting_To_MidCareer_Salary), y = Percent_Change_From_Starting_To_MidCareer_Salary)) +
  geom_bar(stat = "identity") +
  xlab("Undergraduate Major") +
  ylab("Percent Change: Starting to Mid-Career Salary") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


### Conclusions of Majors and Salaries

With these results we can conclude that the best major if you want the highest starting median salary is Physician Assistant at $74,300, the best major if you want the highest mid-career median salary is Chemical Engineering	at $107,000, and the best major for greatest percent growth from starting to mid-career median salary is Math at 103.5%.

# Third Step in Game Plan: Finding the Occupations with the Highest Salaries and Job Security 

We will be using the OccupationSalaries dataset which we found on a website online for the evaluation of this step in the game plan.

First we will extract the table from the website online
```{r}
web_page<-"https://www.bls.gov/oes/current/oes_nat.htm#00-0000"
ListofTables <-
  web_page %>%
  read_html() %>%
  html_nodes( css = "table") %>%
  html_table(fill = TRUE)
```

Next, we will use data scraping to get the data ready for analysis.
```{r}
head(ListofTables[[2]])
OccupationSalaries <- ListofTables[[2]]
OccupationSalaries <- OccupationSalaries %>%
  filter(row_number() > 2) %>%
   rename(Code = `Occupation code`,
          Name =`Occupation title (click on the occupation title to view its profile)`,
          Level = Level,
          Employment = Employment,
          RSE = `Employment RSE`,
          EP1J = `Employment per 1,000 jobs`,
          Median = `Median hourly wage`,
          Mean = `Mean hourly wage`,
          Annual = `Annual mean wage`,
          WRSE = `Mean wage RSE`)
str(OccupationSalaries)
OccupationSalaries <- 
  OccupationSalaries %>%
  mutate(Employment = gsub(pattern = ',', replacement = '', Employment), RSE = gsub(pattern = '%', replacement = '', RSE), Median = gsub(pattern = '\\$', replacement = '', Median), Mean = gsub(pattern = '\\$', replacement = '', Mean), Annual = gsub(pattern = '\\$|,', replacement = '', Annual), WRSE = gsub(pattern = '%', replacement = '', WRSE))

OccupationSalaries<-
  OccupationSalaries %>%
  mutate(Employment = as.numeric(Employment), RSE = as.numeric(RSE),EP1J = as.numeric(EP1J), Median = as.numeric(Median), Mean = as.numeric(Mean), Annual = as.numeric(Annual), WRSE = as.numeric(WRSE))
str(OccupationSalaries)
OccupationSalaries <- na.omit(OccupationSalaries)
OccupationSalaries
```

Now we will join the OccupationSalaries dataset to the OccupationProjection dataset so that we can make a full analysis on Occupations that will provide us with data on salaries as well as job openings. 

```{r}
Occupation_Proj_Salary_Base <-
  OccupationsProjection %>%
  rename(Name = Occupation_Name, Code = Occupation_Code) %>%
  full_join(OccupationSalaries)
Occupation_Proj_Salary <- na.omit(Occupation_Proj_Salary_Base)
Occupation_Proj_Salary
```

## Finding Occupations with the Best Job Security
```{r}
Occupation_Proj_Salary <-
  Occupation_Proj_Salary_Base %>%
  filter(EP1J > 15, Percent_Change > 10,Average_Annual_Openings > 1000) %>%
  arrange(desc(Percent_Change)) %>%
  group_by(Name) %>%
  summarise(Average_Annual_Openings = sum(Average_Annual_Openings)) %>%
  arrange(desc(Average_Annual_Openings)) %>%
  head(50)
Occupation_Proj_Salary
```

In the above analysis we tried to find the jobs that have the best security. We defined security as having a high upward change in job creation, a high average annual job opening and a high EP1J (Employment per 1000 jobs). As an arbitrary value for high we decided on an EP1J of greater than 15, A annual job openings of greater than 1000 and a high upward change in job creation of greater than 10%. We Thus filtered these jobs out of the data-set and the 6 jobs with the highest security are Registered nurses, Retail Salespersons, Customer Service Representatives, Laborers and Freight and General and Operations Managers, Cashiers, and Office Clerks. These jobs are not however, the best in terms of Salary. This is what we will analyse next.

## Visualizations 
```{r}
Occupation_Proj_SalaryGraphic1 <-
  Occupation_Proj_Salary %>%
  ggplot(aes(x = reorder(Name, -Average_Annual_Openings), y = Average_Annual_Openings)) +
  geom_bar(stat = "identity") +
  xlab("Job") +
  ylab("Average Annual Openings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
Occupation_Proj_SalaryGraphic1
```


## Finding Jobs with the Highest Salaries

```{r}
highSalary <-
  Occupation_Proj_Salary_Base %>%
  arrange(desc(Annual)) %>%
  group_by(Name) %>%
  summarise(AnnualSalary = sum(Annual)) %>%
  arrange(desc(AnnualSalary)) %>%
  head(20)
highSalary
```

From the above analysis, it is clear that the jobs with the highest salaries include, Chief Executives, Dentists, Nurse Anesthetists, Computer and Information Systems Managers, Financial Managers and Architectural and Engineering Managers, Lawyers, Marketing Managers, Sales Managers, and Natural Sciences Majors.

## Visualizations 
```{r}
Occupation_Proj_SalaryGraphic2 <-
  highSalary %>%
  ggplot(aes(x = reorder(Name, -AnnualSalary), y = AnnualSalary)) +
  geom_bar(stat = "identity") +
  xlab("Job") +
  ylab("Annual Salary") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
Occupation_Proj_SalaryGraphic2
```


### Conclusions of Jobs and Salaries and Security 

With these results we can conclude that the job with the highest salary is Chief Executives at $9,304,800 annual salary and the job with the best job secruity is Registered Nurses with 210,350 average annual job openings. From this analysis it is clear that the jobs that provide the highest job security do not provide the best security financially. 

# Conclusions for Answering the Research Question 

In conclusion, a student should choose one of the colleges California Institute of Technology (CIT), Massachusetts Institute of Technology (MIT), Harvey Mudd College, Stanford University, or Princeton University. One of the majors you should choose are Physician Assistant, Chemical Engineering, Computer Engineering, Electrical Engineering, or Mechanical Engineering. One of the occupations you should choose are Chief Executive, Dentists, Nurse Anesthetists, Computer and Information Systems Managers, or Financial Managers. We chose these factors because they are the top five for highest starting median salary. 



