---
title: "FinalProjectData"
authors: "Lily Jakielaszek, Allison Denenberg, Ayoolwatomiwa Ogunrombi"
output: html_notebook
---

### Load the Libraries and Datasets
```{r}
# front matter
rm(list = ls())
library(DataComputing)
library(lubridate)
library(tidyverse)
library(rvest)
library(mosaic)
OccupationProjection <- read.csv("https://query.data.world/s/hfgf3loqzzbk34aytghvne5sxg3ljr", header=TRUE, stringsAsFactors=FALSE);

OccupationProjection<-OccupationProjection%>%
  filter(Occupation.Name!="Total, All Occupations")

salaries_by_region = readr::read_csv("salaries-by-region.csv")

salaries_by_college_type = readr::read_csv("salaries-by-college-type.csv")

degrees_that_pay_back = readr::read_csv("degrees-that-pay-back.csv")
```
# Research Question

What occupations and majors by state will have the highest salary and have the greatest future outlook?

## Make New Datasets

These datasets will be the ones that will be used to perform statistics on and joins on. They are renamed from the original datasets and include all the same information except they have R acceptable variable names and the correct type for the their output. 

```{r}
SalariesByRegion <-
  salaries_by_region %>%
  mutate(School_Name = `School Name`,
         Mid_Career_Median_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career Median Salary`)),
         Starting_Median_Salary = as.numeric(gsub('\\$|,', '', `Starting Median Salary`)),
         Mid_Career_10th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 10th Percentile Salary`)),
         Mid_Career_25th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 25th Percentile Salary`)),
         Mid_Career_75th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 75th Percentile Salary`)),
         Mid_Career_90th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 90th Percentile Salary`)),
         Difference = Mid_Career_Median_Salary - Starting_Median_Salary) 
SalariesByRegion <- 
  SalariesByRegion %>%
  select(School_Name, Region, Mid_Career_Median_Salary, Starting_Median_Salary, Mid_Career_10th_Percentile_Salary, Mid_Career_25th_Percentile_Salary, Mid_Career_75th_Percentile_Salary, Mid_Career_90th_Percentile_Salary, Difference)
SalariesByRegion
```
```{r}
SalariesByCollegeType <-
  salaries_by_college_type %>%
  mutate(School_Name = `School Name`,
         School_Type = `School Type`,
         Mid_Career_Median_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career Median Salary`)),
         Starting_Median_Salary = as.numeric(gsub('\\$|,', '', `Starting Median Salary`)),
         Mid_Career_10th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 10th Percentile Salary`)),
         Mid_Career_25th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 25th Percentile Salary`)),
         Mid_Career_75th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 75th Percentile Salary`)),
         Mid_Career_90th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 90th Percentile Salary`)),
         Difference = Mid_Career_Median_Salary - Starting_Median_Salary)
SalariesByCollegeType <-
  SalariesByCollegeType %>%
  select(School_Name, School_Type, Starting_Median_Salary, Mid_Career_Median_Salary, Mid_Career_10th_Percentile_Salary, Mid_Career_25th_Percentile_Salary, Mid_Career_75th_Percentile_Salary, Mid_Career_90th_Percentile_Salary, Difference)
SalariesByCollegeType
```

```{r}
DegreesThatPayBack <-
  degrees_that_pay_back %>%
  mutate(Undergraduate_Major = `Undergraduate Major`,
         Mid_Career_Median_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career Median Salary`)),
         Starting_Median_Salary = as.numeric(gsub('\\$|,', '', `Starting Median Salary`)),
         Mid_Career_10th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 10th Percentile Salary`)),
         Mid_Career_25th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 25th Percentile Salary`)),
         Mid_Career_75th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 75th Percentile Salary`)),
         Mid_Career_90th_Percentile_Salary = as.numeric(gsub('\\$|,', '', `Mid-Career 90th Percentile Salary`)),
         Percent_Change_From_Starting_To_MidCareer_Salary = `Percent change from Starting to Mid-Career Salary`,
         Difference = Mid_Career_Median_Salary - Starting_Median_Salary)
DegreesThatPayBack <-
  DegreesThatPayBack %>%
  select(Undergraduate_Major, Starting_Median_Salary, Mid_Career_Median_Salary, Mid_Career_10th_Percentile_Salary, Mid_Career_25th_Percentile_Salary, Mid_Career_75th_Percentile_Salary, Mid_Career_90th_Percentile_Salary, Percent_Change_From_Starting_To_MidCareer_Salary, Difference)
DegreesThatPayBack
```
```{r}
OccupationProjection <-
  OccupationProjection %>%
  mutate(State_FIPS = State.FIPS,
         Region = Area.Name,
         Occupation_Code = Occupation.Code,
         Occupation_Name = Occupation.Name,
         Base_Year = Base.Year,
         Starting_Median_Salary = Base,
         Projected_Year = Projected.Year,
         Mid_Career_Median_Salary = Projection,
         Percent_Change_From_Starting_To_MidCareer_Salary = Percent.Change,
         Average_Annual_Openings = Average.Annual.Openings) 
OccupationProjection <-
  OccupationProjection %>%
  select(State_FIPS, Region, Occupation_Code, Occupation_Name, Base_Year, Starting_Median_Salary, Projected_Year, Mid_Career_Median_Salary, Change, Percent_Change_From_Starting_To_MidCareer_Salary, Average_Annual_Openings) %>%
  filter(Region != " United States")
OccupationProjection
```
### Inspecting the Revised Datasets 
## OccupationProjection Inspection
```{r}
str(OccupationProjection)
head(OccupationProjection)
tail(OccupationProjection)
nrow(OccupationProjection)
```

## DegreesThatPayBack Inspection
```{r}
str(DegreesThatPayBack)
head(DegreesThatPayBack)
tail(DegreesThatPayBack)
nrow(DegreesThatPayBack)
```

## SalariesByCollegeType Inspection
```{r}
str(SalariesByCollegeType)
head(SalariesByCollegeType)
tail(SalariesByCollegeType)
nrow(SalariesByCollegeType)
```

## SalariesByRegion Inspection
```{r}
str(SalariesByRegion)
head(SalariesByRegion)
tail(SalariesByRegion)
nrow(SalariesByRegion)
```

### Join Datasets Together 
```{r}
Salaries<-salaries_by_college_type%>%
  full_join(salaries_by_region)
na.omit(Salaries)
```

